---
title: "Reproducible Research Peer Assignment 1"
author: "Ann Laska"
date: "March 19, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## Load libraries
library(lattice)
library(knitr)
library(rmarkdown)

## Load activity.csv into a variable
act_org<-read.csv("activity.csv", header = TRUE)
```

###Find sum by day without NA values and plot in a histogram
```{r sumbyday, include=TRUE}
sum_nona<-aggregate(act_org$steps, by = list(act_org$date), na.rm = TRUE, FUN = sum)

## Apply friendly column names
colnames(sum_nona)<-c("Date", "SumSteps")

## Create a histogram of steps per day
hist(sum_nona$SumSteps, xlab = "Sum of Steps", main = "Histogram of Steps")
```

```{r mean_median, include = TRUE}
###Find mean by day without NA values
mean_nona<-aggregate(act_org$steps, by = list(act_org$date), na.rm = TRUE, FUN = mean)

###Apply friendly column names
colnames(mean_nona)<-c("Date", "MeanSteps")

###Find median by day without NA values
med_nona<-aggregate(act_org$steps, by = list(act_org$date), na.rm = TRUE, FUN = median)

## Apply friendly column names
colnames(med_nona)<-c("Date", "MedianSteps")
```
###Print median by day
```{r med_day, include = TRUE}
print(med_nona)
```
###Print mean by day
```{r mean_day, include = TRUE}
print(mean_nona)
```

###Find average by day and interval without NA and plot in a time series line graph
```{r avg_int, include=TRUE}
avg_nona<-aggregate(act_org$steps, by = list(act_org$interval), na.rm = TRUE, FUN = mean)

## Apply friendly column names
colnames(avg_nona)<-c("Interval", "AvgSteps")

## Plot average steps by day/interval
plot(avg_nona$Interval, avg_nona$AvgSteps, type = "l", main = "Avg Steps/Day by Interval", xlab = "Interval", ylab = "Avg Steps")
```

###Find number of NA values in steps
```{r num_na, include=TRUE}
print(sum(is.na(act_org$steps)))
```

```{r imputing, include=TRUE}
## Use the mean interval to fill in the NA values 

## Set the loop variable
i<-1
z<-0
for (i in 1:17568)
  {
    j<-1
    if(is.na(act_org[i,1]))
    {
      for (j in 1:288)
      {
        if(act_org[i,3]==avg_nona[j,1])
        {
          act_org[i,1]<-avg_nona[j,2]
        }
      }
      
    }
    i<-i+1
}
```

##Repeat with imputed values
###Find sum by day and plot in a histogram
```{r sumbyday1, include = TRUE}
## Find sum by day
sum_nona1<-aggregate(act_org$steps, by = list(act_org$date), na.rm = TRUE, FUN = sum)

## Apply friendly column names
colnames(sum_nona1)<-c("Date", "SumSteps")

## Create a histogram of steps per day
hist(sum_nona1$SumSteps, xlab = "Sum of Steps", main = "Histogram of Steps")
```
```{r mean_med1, include=TRUE}
###Find mean by day
mean_nona1<-aggregate(act_org$steps, by = list(act_org$date), na.rm = TRUE, FUN = mean)

###Apply friendly column names
colnames(mean_nona1)<-c("Date", "MeanSteps")

###Find median by day
med_nona1<-aggregate(act_org$steps, by = list(act_org$date), na.rm = TRUE, FUN = median)

###Apply friendly column names
colnames(med_nona1)<-c("Date", "MedianSteps")
```
I chose to use the average per interval to fill in the NA values. Couple of issues:

* I did not distiguish the average on a weekday vs. a weekend for each interval

* I did not throw out any outliers, which could skew the values

The end result is a cleaner bell curve on the histogram because the missing values were normalized by taking the average.
###Print median by day
```{r med_day1, include = TRUE}
print(med_nona1)
```
###Print mean by day
```{r mean_day1, include = TRUE}
print(mean_nona1)
```

###Plot activity weekends vs. weekdays
```{r plot_days, include=TRUE}
## Add factor for Weekday vs. Weekend
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
act_org$date<-as.Date(act_org$date)
act_org$wDay <- factor((weekdays(act_org$date) %in% weekdays1), levels=c(FALSE, TRUE), labels=c('weekend', 'weekday')) 
## Get the average steps
avg_interval_wd<-aggregate(act_org$steps, by = list(act_org$interval, act_org$wDay), FUN = mean)
## Apply friendly names
colnames(avg_interval_wd)<-c("interval", "wDay", "steps")
## Make the panel plot
xyplot(avg_interval_wd$steps~avg_interval_wd$interval|avg_interval_wd$wDay,type="l",group=avg_interval_wd$wDay, layout=c(1,2), xlab = "Interval", ylab = "# of Steps")
```

I would've expected the average on the weekend to be higher. However, the spike in the morning doesn't surprise me much since that's when a lot of people choose to work out.
